---
title: "Interactive Graphs in R using plotly and ggplot2"
author: "Evan Moore"
date: "February 11th, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## Introduction

- It's easy to make interesting and interactive plots in R using packages like `plotly` and `ggplot2`!
- For these examples, we will be using the `diamonds` dataset that comes preloaded with ggplot. 
- View the data dictionary [here](http://ggplot2.tidyverse.org/reference/diamonds.html).

## Getting Started

First, let's load up our required packages and data. 
```{r}
library(ggplot2) #for plotting
library(dplyr) #for data manipulation
library(plotly) #for interactivity
data(diamonds)
```

## Diamonds 

What variables are included in the `diamond` dataset, and what are their types? Getting a good sense of the variables included in your data is the first step to making great visualizations! 
```{r}
glimpse(diamonds)
```

## Initial Graphing Ideas

Let's try a scatter plot between carat and price, adding in a smoothed regression line to help us see the relationship between these variables. `ggplot2` makes it very easy to "stack" aethetic layers in this way.
```{r}
diamonds <- diamonds[sample(nrow(diamonds), 5000), ]
p <- ggplot(diamonds, aes(x = carat, y = price)) + 
  geom_point() + 
  geom_smooth()
```

## The Plot

Let's check out the plot we just created. 
```{r, echo = F}
p
```

## Expanding on our Aesthetics

The plot looks nice, but seems to be missing something...color! Notice how easy it is to add an additional parameter to the plot we just created. 
```{r}
p2 <- p + aes(col = color)
```

## Graphing with Color

Looks much better now. However, the color argument has the additional effect of creating unique regression lines for each color, making the plot more cluttered and hard to interpret as a result.
```{r, echo = F}
p2
```

## Plotly to the Rescue

Here's where `plotly` comes in! The `ggplotly()` function can be applied to (almost) any `ggplot2` object in order to make it interactive, with additional features like toggling which data to show and ability to zoom in and hover on specific data points.

## Interactive ggplot 1
```{r}
ggplotly(p2)
```

## GGplot 

Attempting to make similar graphs in ggplot requires extra effort, without the added customizability. 
```{r}
p <- diamonds %>% 
  filter(color == "D" | color == "J") %>% 
  ggplot(aes(x = carat, y = price, col = color)) + 
  geom_point() + 
  geom_smooth()
```

## ggplot Attempt
```{r, echo = F}
p
```

## More Plotly Examples

We can try a similar plot to before, adding in an additional variable to facet by as well as changing the type of the regression line from smoothed to linear. 
```{r}
p<- ggplot(diamonds, aes(x =carat, y = price, col = color)) + 
           geom_point() + 
           geom_smooth(method="lm") + 
           facet_grid(~clarity)
```

## Interactive ggplot 2
```{r, echo = F}
ggplotly(p)
```

## Interactive ggplot 3
Let's try the same plot, but swap what we are faceting and coloring by.
```{r, echo = F}
ggplotly(ggplot(diamonds, aes(x =carat, y = price, col = clarity)) + 
    geom_point() + 
    geom_smooth(method="lm") + 
    facet_grid(~color))
```

## Plotly Graphing
`plotly` also offers many options for creating graphs through their own syntax. Check out [Plotly's website](https://plot.ly/r/) to see the full list of options (as well as code for each). The syntax is generally simple! For example, let's try making a heatmap.
```{r}
p <- plot_ly(data = diamonds, 
    x = ~cut, 
    y = ~color, 
    z = ~price, 
    type="heatmap")
```

## Heatmap
```{r, echo = F}
p
```

## More Options
Another fun and easy option is the 3d scatterplot. Be aware, however, that it can be easy to create ugly or nonsenseical graphs in such a way, especially in 3 dimensions...
```{r}
#bad example
p <- plot_ly(data = diamonds, 
    x = ~cut, 
    y = ~color, 
    z = ~price, 
    type="scatter3d")
```

## (Bad) 3D Scatterplot
```{r, echo = F}
p
```

## (Slightly Better) 3D Scatterplot
```{r,echo = F}
plot_ly(data = diamonds, x = ~carat, y = ~color, z = ~price, color = ~clarity, type = "scatter3d")
```

## ggplot-like Graphs
Plotly can also do graphs similar to ggplot's style, with the added benefit of offering customizable hover-text and other interactivity options. Check this out! 
```{r}
p <- plot_ly(
  diamonds, x = ~carat, y = ~price,
  # Hover text:
  text = ~paste("Price: ", price, '$<br>Cut:', cut),
  color = ~carat, size = ~carat
)
```

## Plotly 2D scatter plot
```{r, echo =F}
p
```

## Conclusion
We've only scratched the surface of the possibilities `plotly` offers in R, but we hope you enjoyed learned about the basics of the packages, as well as its easy integrability with standard R plotting libraries `ggplot2`. 